### 2022-06-17 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@fa58cdffbc410db83637c21cff76d58ca3bd6ac3 🚀

\noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of
bat-betacoronaviruses spillover risk}}}

@@

Based on the previous result, we extracted the risk component from the
composite map (see Methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.
However, this maps the potential risk, which must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et al.
2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in
fig.~\ref{fig:compound}. Most of South America and Europe are at low
risk, as although densely populated, settlements tend to be in areas
with lower potential risk. However, this mapping reveals that South-East
Asia, the Indian subcontinent, and parts of sub-Saharan Africa, are at
high risk due to the overlap between built areas and bat communities
representing more opportunities for cross-species transmission of
betacoronaviruses. In looking for the origins of SARS in China, Xu et
al. (2004) present serological evidence that strongest human-animal
contact results in higher risk of virus exposure, regardless of the
animal species, but that different types of contact had different
impacts. Ideally, finer-grained information about human activity (rather
than human presence through anthropisation) could allow to partition
this risk further.

\begin{figure}
\hypertarget{fig:compound}{%
\centering
\includegraphics{figures/risk_compounded.png}
\caption{Overlap of the percent of each pixel occupied by urbanized
structures, representing the degree of settlement, on the risk map.
Darker pixels correspond to more risk, in that the GMTC-derived risk of
fig.~\ref{fig:risk} is high \emph{and} the pixel is densely occupied by
human populations. This approach increases the relative risk of several
regions in Africa, and highlights the risk in India, southeast China,
and the arabic peninsula where areas of high to moderate risk overlap
with areas of denser population.}\label{fig:compound}
}
\end{figure}

@@

Our study focuses largely on the biogeography of hosts. Yet, we know
that viruses with high host plasticity, that is, the ability of a given
virus to adapt to various taxonomic orders and ecological groups
(Kreuder Johnson et al. 2015), are more likely to amplify viral
spillover, followed by secondary human-to-human transmission, and
geographical spread (Hazarie et al. 2021). High viral host plasticity is
an especially important trait for RNA viruses like betacoronaviruses
(Haddad et al. 2021). Indeed, our analysis of viral sequences reveals
that Latin America is a hotspot of viral distinctiveness, suggesting
that this part of the bats-betacoronavirus system may be undergoing
independent evolutionary dynamics (related species sharing viruses that
are different from the rest of the global pool). The other hotspot of
viral distinctiveness is S.E. Asia, in which richness is high but
sharing is low; this suggests a different type of evolutionary dynamics
(unrelated viruses coevolving with evolutionarily distinct hosts,
generating high diversity locally). Both of these areas should be
priority areas for sampling, especially since Becker et al. (2022)
advance that they harbor undiscovered hosts of beta-coronaviruses. This
diversity of hosts, and the mechanisms by which the exchange of viruses
occurs between species, is largely affected by the local environmental
conditions and environmental change.

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human health (Letko et al.
2020, Van Brussel and Holmes 2022). Chiropterans emerged around 64
million years ago and are one of the most diverse mammalian orders, with
an estimated richness of more than 1400 species (Peixoto et al. 2018,
Simmons and Cirranello 2020). They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to human (Kasso and Balakrishnan
2013). For example, bats are an essential component of many
seed-dispersal networks (Mello et al. 2011). Over two-thirds of bats are
know to be either obligate or facultative insectivores, therefore
playing an important role in the regulation of insect pests that can
affect crops (Williams-Guillén et al. 2008, Voigt and Kingston 2016),
and vectors of pathogens that put a risk on human health (Gonsalves et
al. 2013a, b). Because bats are globally distributed and have a long
evolutionary history, phylogeographic and biogeographic approaches are
required to shed light on the contemporary distribution of
coevolutionary processes between bats and the pathogens they host. Not
all areas in which bats, viruses, and human are co-occuring are facing a
risk of spillover towards human populations, and the areas in which this
risk exist may not be facing risks of the same nature and magnitude.

@@

@@

@@

@@

\leavevmode\hypertarget{ref-Muylaert2021PreFut}{}%
Muylaert, R. L. et al. 2021. Present and future distribution of bat
hosts of sarbecoviruses: Implications for conservation and public
health.

@@

>>>>

\noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of
\emph{Betacoronavirus} spillover risk from bat hosts}}}

@@

Based on the previous result, we extracted the risk component from the
composite map (see Methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}. As
this map represents the potential risk, it must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et al.
2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in
fig.~\ref{fig:compound}. Most of South America and Europe are at
comparatively lower risk, as although densely populated, settlements
tend to be in areas with lower potential risk. Regions like Malaysia and
the North coast of Australia have a high risk component, but should
represent a relatively lower effective risk due to low human density.
However, this mapping reveals that South-East Asia, the Indian
subcontinent, and parts of sub-Saharan Africa, are at high risk due to
the overlap between built areas and bat communities representing more
opportunities for cross-species transmission of betacoronaviruses. In
looking for the origins of SARS in China, Xu et al. (2004) present
serological evidence that strongest human-animal contact results in
higher risk of virus exposure, regardless of the animal species, but
that different types of contact had different impacts. Ideally,
finer-grained information about human activity (rather than human
presence through anthropisation) could allow to partition this risk
further, alebit at the cost of more hypotheses required to estimate the
amount of risk represented by each activity. Our map of risk overlays
with recent results from Cohen et al. (2022) -- areas of purported high
risk/diversitifcation potential (Madagascar, South-America) overlay with
sampling gaps for \emph{Betacoronavirus}, stressing the need for
spatially targeted monitoring and discovery.

\begin{figure}
\hypertarget{fig:compound}{%
\centering
\includegraphics{figures/risk_compounded.png}
\caption{Overlap of the percent of each pixel occupied by urbanized
structures, representing the degree of settlement, on the spillover risk
map (where the risk comes only from wildlife, and ignores multi-hosts
chains of transmissions including non-bats hosts). Darker pixels
correspond to more risk, in that the GMTC-derived risk of
fig.~\ref{fig:risk} is high \emph{and} the pixel is densely occupied by
human populations. This approach increases the relative risk of several
regions in Africa, and highlights the risk in India, southeast China,
and the Arabian peninsula where areas of high to moderate risk overlap
with areas of denser population.}\label{fig:compound}
}
\end{figure}

@@

Our study focuses largely on the biogeography of hosts. Yet, we know
that viruses with high host plasticity, that is, the ability of a given
virus to adapt to various taxonomic orders and ecological groups
(Kreuder Johnson et al. 2015), are more likely to amplify viral
spillover, followed by secondary human-to-human transmission, and
geographical spread (Hazarie et al. 2021). High viral host plasticity is
an especially important trait for RNA viruses like betacoronaviruses
(Haddad et al. 2021). Indeed, our analysis of viral sequences reveals
that Latin America is a hotspot of viral distinctiveness, suggesting
that this part of the bats-betacoronaviruses complex may be undergoing
independent evolutionary dynamics (related species sharing viruses that
are different from the rest of the global pool). The other hotspot of
viral distinctiveness is S.E. Asia, in which richness is high but
sharing is low; this suggests a different type of evolutionary dynamics
(unrelated viruses coevolving with evolutionarily distinct hosts,
generating high diversity locally, see \emph{e.g.} Latinne et al. 2020).
Both of these areas should be priority areas for sampling, especially
since Becker et al. (2022) advance that they harbor undiscovered hosts
of beta-coronaviruses. This diversity of hosts, and the mechanisms by
which the exchange of viruses occurs between species, is largely
affected by the local environmental conditions and environmental change.

Bats are important reservoir hosts for different classes of
microorganisms, many of which a threat to human health (Letko et al.
2020, Van Brussel and Holmes 2022). Chiropterans emerged around 64
million years ago and are one of the most diverse mammalian orders, with
an estimated richness of more than 1400 species (Peixoto et al. 2018,
Simmons and Cirranello 2020). They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to human (Kasso and Balakrishnan
2013). For example, bats are an essential component of many
seed-dispersal networks (Mello et al. 2011). Over two-thirds of bats are
know to be either obligate or facultative insectivores, therefore
actively contributing for agricultural pest control (Williams-Guillén et
al. 2008, Voigt and Kingston 2016), and vectors of pathogens that put a
risk on human health (Gonsalves et al. 2013a, b). Because bats are
globally distributed and have a long evolutionary history,
phylogeographic and biogeographic approaches are required to shed light
on the contemporary distribution of coevolutionary processes between
bats and the pathogens they host. Not all areas in which bats, viruses,
and human are co-occuring are facing a risk of spillover towards human
populations, and the areas in which this risk exist may not be facing
risks of the same nature and magnitude.

@@

\leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
Kalyaanamoorthy, S. et al. 2017. ModelFinder: Fast model selection for
accurate phylogenetic estimates. - Nature Methods 14: 587--589.

@@

\leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
Katoh, K. and Standley, D. M. 2013. MAFFT Multiple Sequence Alignment
Software Version 7: Improvements in Performance and Usability. -
Molecular Biology and Evolution 30: 772--780.

@@

\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
Latinne, A. et al. 2020. Origin and cross-species transmission of bat
coronaviruses in China. - bioRxiv: The Preprint Server for Biology:
2020.05.31.116061.

@@

\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
Muylaert, R. L. et al. 2022. Present and future distribution of bat
hosts of sarbecoviruses: Implications for conservation and public
health. - Proceedings of the Royal Society B: Biological Sciences 289:
20220397.

\leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
Nguyen, L.-T. et al. 2015. IQ-TREE: A Fast and Effective Stochastic
Algorithm for Estimating Maximum-Likelihood Phylogenies. - Molecular
Biology and Evolution 32: 268--274.

@@

\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
Villalobos, F. and Arita, H. T. 2010. The diversity field of New World
leaf-nosed bats (Phyllostomidae). - Global Ecology and Biogeography 19:
200--211.
